Meta-analysis

### 2020-06-22 04_MetaAnalysis.Rmd
Wordsmithing and flow

First lets compare models with and without a log transformed response variable to evaluate if the log transformed model better fits our assumptions of normality.  
Next we will test whether including `spray_management` as a fixed effect variable significantly accounts for enough variation to be included in the model.  
Finally we will test whether including `spray_management`  as a random slope significantly accounts for enough variation to be used in the model.

@@

This linear mixed effect model shows indicates:

@@

```{r remove_NA_MSE}
slimmer_PM_dat <- 
   slimmer_PM_dat[!is.na(slimmer_PM_dat$Y_Msquare),]

@@

@@

There seems like little difference between the treatments, with exception to Late_plus. 
Let's do this plot again, but lets look at the proportional mean difference between the treatments and the no spray control for each study.  
This should reduce variation in yield due between trials.

@@

Using the mean difference in the treatment effect seems to show the impact of each treatment better than just raw yield.
We know there is a good deal of variability in our studies and between them. 
Mungbean produces variable yields between seasons so we should use a response that highlights the difference in the treatment effects that we are interested in and reduce the variability.
Lets calculate the standardised mean difference for each treatment to reduce the variability in the meta-analysis.

@@

@@

The sample variance is a similar formula except we use $x$ instead of $y$.  
$S_{Var} = \frac{\sum_{x_1}^{x_n}(x_i - \overline{x})^2}{n - 1}$  
$x$ being each observation within the experiment and $n$ being the number of samples in the whole trial.
Therefore: 
$S_{Var} = \frac{\frac{\sum_{y_1}^{y_n}(y_i - \overline{y})^2}{n_y - 1}}{n_{x}}$  
Where $n_x$ is the number of samples within each treatment.

@@

### Caluculate standardised mean differences
We may wish to undertake the meta-analysis using standardised mean differences. 
Lets calculate them and add them to the dataframe.
```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for(T_ref in slimmer_PM_dat$trial_ref){

@@

We can use the same sample variance because if you add or subtract from a random variable the variance does not change.

Lets simplify our data
```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
      spray_management,
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )))%>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      vi,
      vi2
   )
```

## Meta-analysis  
### metafor package  
Let's load the `metafor` package we are using to analyse the data, then rearrange the factors we want to examine by placing the control treatment first. This way all treatments will be compared to the no spray controls.  
Next we are log transforming the grain yield and calculating the variance from the trial mean squares.  
Finally we will assign factor classes to the main variables in the meta-analysis. Note that variable trial is a combination of:

   * Trial identifier

   * Trial year

   * Trial location

   * Host genotype

   * Trial row spacing

Lets inspect the data to determine if we need to transform the response variable

@@

Lets undertake the meta-analysis, we are using the `spray_management` variable as a moderator and a interactive term to the `trial` variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.
```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management -1,
   method = "ML",
   random = ~ spray_management|trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero, which in this case is the mean of the no spray control. 
However the other treatments are significantly different to zero.
The Qm [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, show the moderators significantly influence the model ($Q_M = 85.232, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010]. 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 40994.51, df = 105, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

Lets see if we can improve this model by including some other random effects and improving the random effect structure in the model.

```{r}

@@

AIC(PM_mv_St)
anova(PM_mv, PM_mv_St)
```
The simpler `PM_mv_St` model explains significantly less variation as the more complicated model `PM_mv`. So we will keep the original model.

@@

### Profile plots  
Lets inspect the profile plots to ensure the model is not over-fitted. 
We expect to see the estimate align with the peak of the curve. 
Also that the shape of the line is a curve.  
Caution! this will take some time to run.

>>>>

First let's compare models with and without a log transformed response variable to evaluate if the log transformed model better fits our assumptions of normality.  
Next we will test whether including `spray_management` as a fixed effect variable significantly accounts for enough variation to be included in the model.  
Finally we will test whether including `spray_management`  as a random slope significantly accounts for enough variation to be used in the model.

@@

This linear mixed effect model shows:

@@

```{r remove_yieldErrorNAs}
slimmer_PM_dat <- 
   slimmer_PM_dat[!is.na(slimmer_PM_dat$yield_error),]

@@

Let's visualise the spread of data in each treatments with box-plots.

@@

There seems like little difference between the treatments, with exception to Late_plus. 
Let's do this plot again, but let's look at the proportional mean difference between the treatments and the no spray control for each study.  
This should reduce variation in yield due between trials.

@@

Using the mean difference in the treatment effect seems to show the impact of each treatment better than just raw yield and reduces the variability in the treatments.
We know there is a good deal of variability in our studies and between them. 
Mungbean produces variable yields between seasons so we should use a response that highlights the difference in the treatment effects that we are interested in and reduce the variability.
Let's calculate the standardised mean difference for each treatment to reduce the variability in the meta-analysis.

@@

Currently our data frame has the sample variance calculated earlier, however we can approximate it using the mean squared error.
Let's show the rational for our calculation and then add a secondary variance column (`vi2`)

@@

The sample variance is a similar formula except we use $x$ instead of $y$.  
$S_{Var} = \frac{\sum_{x_1}^{x_n}(x_i - \overline{x})^2}{n - 1}$  
$x$ being each observation within the experiment and $n$ being the number of samples in the whole trial.
Therefore: 
$S_{Var} = \frac{1}{n_{x}} \frac{\sum_{y_1}^{y_n}(y_i - \overline{y})^2}{n_y - 1} $  
Where $n_x$ is the number of samples within each treatment.

@@

### Calculate standardised mean differences

Let's calculate standardised mean differences and add them to the data frame.
```{r StandardisedMeanDifference}
slimmer_PM_dat$pooledSD <- NA
slimmer_PM_dat$vi_C <- NA
for(T_ref in slimmer_PM_dat$trial_ref){

@@

We can use the same sample variance with the standardised mean difference.
This is because if you add or subtract from a random variable, the variance does not change.

Let's simplify our data
```{r metafor_organisation}
dat1 <-
   slimmer_PM_dat %>%
   filter(fungicide_ai != "control") %>% #remove controls from the data
   mutate(spray_management =
             factor(
      spray_management,
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )))%>%
   select(
      trial,
      trial_ref,
      location,
      year,
      row_spacing,
      host_genotype ,
      spray_management,
      fungicide_ai,
      D_pres,
      grain_yield.t.ha,
      grain_SMD,
      vi,
      vi2
   )
```

## Meta-analysis  
### metafor package  
Let's inspect the data to determine if we need to transform the response variable

@@

Let's undertake the meta-analysis using the package `metafor`.
Ee are using the `spray_management` variable as a moderator and a interactive term to the `trial` random variable.
Because we want to know the difference between all treatments and we have no reference treatment, we will remove the intercept.
```{r Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_SMD,
   vi, 
   mods = ~ spray_management -1,
   method = "ML",
   random = ~ spray_management|trial,
   struct = "UN",
   data = dat1
)
summary(PM_mv)
```
In this result we can see that the `Early` treatment is not significantly different to zero, which in this case is the mean of the no spray control. 
However the other treatments are significantly different to zero.
The Qm [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, show the moderators significantly influence the model ($Q_M = 85.232, df = 5, p < 0.0001$) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010]. 
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E = 47501.48, df = 111, p < 0.0001$) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

Let's see if we can improve this model by including some other random effects and improving the random effect structure in the model.

```{r rma_noRandomInteractiveTerm}

@@

AIC(PM_mv_St)
anova(PM_mv, PM_mv_St)
```
The simpler `PM_mv_St` model explains significantly less variation as the more complicated model `PM_mv`. So we will keep the original model `PM_mv`.

@@

### Profile plots  
Let's inspect the profile plots to ensure the model is not over-fitted. 
We expect to see the estimate align with the peak of the curve. 
Also that the shape of the line is a curve.  
Caution! this will take some time to run.
